VariantSpark: Population Scale Clustering of Genotype Information

### 2015-11-27 doc/publications/journal2014/paper.tex
fixed typos in r/python/adm/pgp

\subsection*{Python Implementation}
Our Python implementation reads in lines from a VCF file as tab-separated values using {\sc DataFrame.read\textunderscore{}csv}, and stores the data in a pandas DataFrame (\url{http://pandas.pydata.org}). The column headings are the individual IDs and the row headings are the allele locations.
We the first 9 columns and convert the remaining allele strings to numeric values.
We convert the DataFrame to a matrix with {\sc .as\textunderscore{}matrix()} and cluster the matrix using sci-kit learn (\url{http://scikit-learn.org/stable/}).

@@

\subsection*{Personal Genome Project Data}
Genome data was downloaded from the Personal Genome Project Data (PGP) website (\url{https://my.pgp-hms.org/public_genetic_data}). Genotype data from 23andMe microarray platforms was sorted by genome 
build. Shell scripts were written to convert NCBI build 36 files to BED format, update to build 37 using UCSC liftOver (\url{https://genome.ucsc.edu/cgi-bin/hgLiftOver}) and 
then convert back to 23andMe format. The 23andMe files were then converted to VCF format using code obtained from the Broad Institute (\url{http://apol1.blogspot.com.au/2013/08/impute-apoe-and-apol1-with-23andme.html}). 
Where individuals had genotype data for both genome builds 36 and 37, only the latter was used. Finally, individual VCF files were combined for clustering into one 
file using VCFtools \cite{Danecek2011Variant}.

@@

%% Use of \listoftables is discouraged.
%%
\section*{Tables}
\label{fivewaycomparison}
\begin{table}[h!]
\caption{The resource consumption of the six compared methods as well as the accuracy measured as \ARI{} on chromosome 22.}
      \begin{tabular}{|c|ccc|ccc|c|c|}
        \hline
           Tool &  \multicolumn{3}{ c |}{Pre-processing} & \multicolumn{3}{ c |}{Clustering} & Accuracy \\
& threads & memory & time  &threads & memory & time  & \\
  \hline
\variantSpark{}	& 8	& 32	& 2min 58sec	& 8	& 32	& 1min 20sec	& 0.84	\\ 
{\sc ADAM}		& 8	& 32	& 12min 48sec	& 8	& 32	& 1min 52sec	& 0.84	\\
Hadoop		& 8	& 32	& 14min 22sec	& 8	& 32	& 14min 23sec	& 0.84	\\
R			& 1	& 32	& 34min 30sec	& 8	& 32	& 7min 25sec	& 0.84	\\
Python		& 1	& 32	& 92min		& 8	& 32	& 11min 29sec	& 0.84	\\
{\sc Admixture}	& 1	& 32	& 10min 8sec	& 8	& 32 & 8min 19sec	& 0.25	\\
  \hline
      \end{tabular}
\end{table}

>>>>

\subsection*{Python Implementation}
Our Python implementation reads in lines from a VCF file as tab-separated values using {\sc DataFrame.read\textunderscore{}csv}, and stores the data in a pandas DataFrame (\url{http://pandas.pydata.org}). The column headings are the individual IDs and the row headings are the allele locations.
We remove the first 9 columns and convert the remaining allele strings to numeric values by applying our {\sc Hamming} function to the DataFrame.
We then convert the DataFrame to a matrix with {\sc .as\textunderscore{}matrix()} and cluster the matrix using `scikit-learn' (\url{http://scikit-learn.org/stable/}).

@@

\subsection*{Personal Genome Project Data}
We acquired genomic data from the Personal Genome Project (PGP) website (\url{https://my.pgp-hms.org/public_genetic_data}). We sorted the genotype data from 23andMe microarray platforms by genome 
build. Using custom shell scripts, we convert NCBI build 36 files to BED format, update to build 37 using UCSC liftOver (\url{https://genome.ucsc.edu/cgi-bin/hgLiftOver}) and 
then convert back to 23andMe format. We then convert the 23andMe files to VCF format using code obtained from the Broad Institute (\url{http://apol1.blogspot.com.au/2013/08/impute-apoe-and-apol1-with-23andme.html}). 
Where individuals had genotype data for both genome builds 36 and 37, only the latter was used. Finally, we combined the individual VCF files into one 
file for clustering, using VCFtools \cite{Danecek2011Variant}.

@@

%% Use of \listoftables is discouraged.
%%
\section*{Tables}
\label{fivewaycomparison}
\begin{table}[h!]
\caption{The resource consumption of the six compared methods as well as the accuracy measured as \ARI{} on chromosome 22.}
      \begin{tabular}{|c|ccc|ccc|c|c|}
        \hline
           Tool &  \multicolumn{3}{ c |}{Pre-processing} & \multicolumn{3}{ c |}{Clustering} & Accuracy \\
& threads & memory & time  &threads & memory & time  & \\
  \hline
\variantSpark{}	& 8	& 32	& 2min 58sec	& 8	& 32	& 1min 20sec	& 0.84	\\ 
{\sc ADAM}		& 8	& 32	& 12min 48sec	& 8	& 32	& 1min 52sec	& 0.84	\\
Hadoop		& 8	& 32	& 14min 22sec	& 8	& 32	& 14min 23sec	& 0.84	\\
R			& 1	& 32	& 34min 30sec	& 8	& 32	& 7min 25sec	& 0.84	\\
Python		& 1	& 32	& 34min 15sec	& 8	& 32	& 11min 29sec	& 0.84	\\
{\sc Admixture}	& 1	& 32	& 10min 8sec	& 8	& 32 & 8min 19sec	& 0.25	\\
  \hline
      \end{tabular}
\end{table}
